---
title: "Assignment 6"
author: "Hanna Matera"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
getwd()

```

## task 1

Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.

```{r cars}
summary(cars)
#loading packages 
library(tidyverse)
library(ggplot2)
library(gganimate)
library(gapminder)



#defensive function calculating GDP in Denmark
#year and country arguments are created in order to be able to inspect specific years and countries from Gapminder data set 

calcGDP <- function(gapminder, year=NULL, country=NULL) {
  if(!is.null(year)) {
    gapminder <- gapminder[gapminder$year %in% year, ]
  }
  if (!is.null(country)) {
    gapminder <- gapminder[gapminder$country %in% country,]
  }
  gdp <- gapminder$pop * gapminder$gdpPercap

  new <- cbind(gapminder, gdp=gdp)
  return(new)
}


#inspecting GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.
head(calcGDP(gapminder, year=c(1967, 1977, 1987, 1997, 2007, 2017),country = "Denmark"))
# N.B.2017 data not present in the gapminder data set 


```

## task 2

Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}

#finding the countries which name starts with B (unique() displayes all members of a given ctegory, here countries), adding ^B and grep() displays all countries from the gapminder data that start with B 

B_countries <- grep("^B", unique(gapminder$country), value=TRUE)
B_countries


gapminder <- as.data.frame(gapminder) 


# I am creating a function that computes average life expectancies across all years for all countries that start with a letter B. Then I am creating a formula that extracts a specific name ofthe coutry and matches it with its average life expectancy. The output is then put in a sentence and printed out. 

for (iCountry in B_countries) {
  tmp <- mean(gapminder[gapminder$country == iCountry, "lifeExp"])
  if(tmp >= 70){
    print(paste("Average Life Expectancy in", iCountry, "is equal to or larger than")) }  
  else if (tmp>=50) {
   print(paste("Average Life Expectancy in", iCountry, "is between 50 and 70 years"))}
  else { print(paste("Average Life Expectancy in", iCountry, "is smaller than 50"))
  }}



```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
