---
title: "Assignment 4"
author: "Hanna Matera"
date: "10/4/2020"
output: html_document
---

```{r setup, include=FALSE}

#First of all, I am defining my working directory 
setwd("~/Desktop/semester 5/CDS/Projects/data")
getwd()


``` 

#Task 1:  Use R to figure out how many elements in the vector below are greater than 2 

```{r}
#loading rooms vector 
rooms <- c(1, 2, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, NA)

#removing missing values
rooms<-rooms[!is.na(rooms)]
print(rooms)

#selecting rooms that are bigger than 2 
big_rooms <- sum(rooms > 2)
```

### Exercise 2: 

What is the result of running median() function on the above 'rooms' vector? (again, best remove the NAs)

```{r}
#running median function on rooms 
median(rooms)

#the median value of 'rooms' is 1.5

```
## Exercise 3: 
Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing a) the script you used to create the object, b) the 'interviews' object in the Environment and the c) structure of your R project in the bottom right Files pane. Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS). Place here the URL leading to the screenshot in your repository.

```{r}
#loading tidyverse package 
library(tidyverse)

#downloading data 
download.file("https://ndownloader.figshare.com/files/11492171",
              "data/SAFI_clean.csv", mode = "wb")

#loading data and putting it in a adata frame using read.csv
interview <- read.csv("SAFI_clean.csv",header=TRUE,sep=",")



knitr::opts_chunk$set(echo = FALSE)
```

